---
title: "Higher Information and Higher Trust in Government ..."
subtitle: "My subtitle if needed"
author: 
  - Talia Fabregas
thanks: "Code and data are available at: [https://github.com/taliafabs/US-Midterms-2022.git](https://github.com/taliafabs/US-Midterms-2022.git)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence. Key Trump 2024 voting blocs, including young men without a college education were less likely to vote in the 2022 midterms when he was not on the ballot compared to 2020"
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(janitor)
library(dplyr)
library(ggplot2)
library(arrow)
library(rstanarm)
library(modelsummary)
library(kableExtra)
library(RColorBrewer)

# load the analysis data
setwd(here::here())
ces2022_analysis_data <- read_parquet("data/02-analysis_data/ces2022_analysis_data.parquet")
```

# Introduction

Overview paragraph

Estimand paragraph
This study aims to examine the effects of 2020 presidential vote preference, age, highest level of education completed, trust in the federal government, knowledge of which party holds a majority in Congress, and interest in politics on the likelihood that an individual who voted for Donald Trump or Joe Biden in the 2020 U.S. presidential election would also vote in the 2022 U.S. midterm election. The estimand, which can never be known with complete certainty, is the true effect of presidential vote preference, age, education, trust, civic knowledge, and interest in politics on the likelihood that an individual who voted in a presidential election will also vote in the subsequent midterm election [@tswd] .


Results paragraph
Based on the results of applying logistic regression models to predict voter turnout and vote preference, I estimate that ...

Why it matters paragraph
In the 2024 election, Democratic Senate candidates (both incumbent and non-incumbent) were elected in states that President-Elect Trump carried: Michigan, Wisconsin, Nevada, and Arizona. In Nevada and Wisconsin, Senators Jackie Rosen and Tammy Baldwin

Telegraphing paragraph: 
The remainder of this paper is structured as follows. @sec-data contains an overview of the survey data set from @ces2022 that was used, visualizations of different variables, and summary statistics. @sec-model contains the logistic regression models used to predict the probability that an individual who voted for either Trump or Biden in the 2020 presidential election would also vote in the 2022 midterm election and the probability that an individual would support Trump over Biden in 2020. @sec-results contains tables and data visualizations that present results about what characteristics made individuals who voted in the 2020 presidential election more likely to also vote in the 2022 midterm election and what characteristics make individuals more likely to support Trump. @sec-discussion contains detailed discussiona about the results presented in @sec-results ...

# Data {#sec-data}

## Overview

The survey data set from @ces2022 contains ...

We use the statistical programming language R [@citeR].... Our data [@ces2022].... Following @tswd, we consider...

Overview text

## Measurement

The primary measurement task is to capture voter characteristics and translate it into conclusions and predictions about voting behavior in U.S. presidential, senate, and congressional elections. During every midterm and presidential election cycle, teams of pollsters work together to conduct surveys that measure public opinion, voter demographics, and candidate preferences. Each survey response is an individual voter's preferences, which pollsters collect to build the Comprehensive Election Study (CES) dataset for each election cycle.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-fig1), from ...

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

```{r}
#| echo: false
#| message: false
#| warning: false
#| eval: true
#| label: tbl-table1
#| tbl-cap: Biden 2020 Voters Voted in 2022 midterms slightly more than Trump 2020 voters. 73.26% of respondents who voted for Trump in 2020 voted in 2022, compared to 74.67% of respondents who voted for Biden.

# turnout_by_preference <- ces2022_analysis_data %>%
#   group_by(gender, age_bracket, educ, trust_state, know) 

# kable(turnout_by_preference)
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| eval: true

ces2022_race <- ces2022_analysis_data |>
  filter(gender == "Female" | gender == "Male")

ces2022_race |>
  ggplot(aes(x = race, fill = presvote2020)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(gender)) +
  theme_minimal() +
  labs(
    title = "White Voters Were More Likely to Support Trump in 2020",
    x = "Race",
    y = "Number of respondents",
    fill = "2020 presidential vote"
  ) +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

ces2022 <-  ces2022_analysis_data |>
  filter(gender == "Male" | gender == "Female") |>
  mutate(educ = factor(educ, levels=c("No HS", "High school graduate", "Some college", "2-year college degree",
                                      "4-year college degree", "Post-grad")))

ces2022 |>
  ggplot(aes(x = educ, fill = presvote2020)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(gender)) +
  theme_minimal() +
  labs(
    title = "College-Educated Voters are Less Likely to Support Trump",
    x = "Highest education",
    y = "Number of respondents",
    fill = "2020 presidential vote"
  ) +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
trust_data <- ces2022_analysis_data %>%
  group_by(know_us_house) %>%
  summarise(
    total_respondents = n(),                     # Total number of respondents
    num_voted = sum(voted_in_2022),              # Total number that voted
    .groups = "drop"                             # Ensure the result is ungrouped
  ) %>%
  mutate(
    pct_voted = round(100*(num_voted/total_respondents),2),
    prop_voted = round(num_voted/total_respondents, 2)
  )

# Trust among Trump and Biden voters
trust_data_by_candidate <- ces2022_analysis_data %>%
  group_by(trustfed, presvote2020) %>%
  summarise(
    total_respondents = n(),                    
    .groups = "drop"                            
  ) 

trust_data_by_candidate$trustfed <- factor(trust_data_by_candidate$trustfed, 
                                           levels = rev(c("A great deal", "A fair amount", "Not very much", "None at all")))

ggplot(trust_data_by_candidate, aes(x = trustfed, y = total_respondents, fill = presvote2020)) +
  geom_bar(stat = "identity", position = "dodge") +   # Position bars side by side
  labs(
    title = "Trump Voters are Less Likely to Trust Government",
    x = "Trust in Federal Government",
    y = "Total Number of Respondents",
    fill = "2020 Presidential Vote"
  ) +
 scale_fill_brewer(palette = "Set1") + 
  theme_minimal() +
  theme(
    legend.position = "right",          # Position legend on top
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis text for readability
  ) +
  coord_flip()
```

```{r}
#| label: fig-fig1
#| fig-cap: More than half of 18-29 year old respondents who voted for Trump in 2020 presidential election did not vote in the 2022 midterm election.
#| echo: false
#| warning: false
#| message: false

# graph the percentage of 2020 voters who also voted in 2022
# age_data <- ces2022_analysis_data %>%
#   group_by(age_bracket, presvote2020) %>%
#   num_voted = (sum(voted_in_2022))
  # summarise(n = sum(voted_in_2022)) %>%
  # mutate(turnout = n / sum(n))

# age_data <- ces2022_analysis_data %>%
#   group_by(age_bracket, presvote2020) %>%
#   summarise(
#     total_respondents = n(),                     # Total number of respondents
#     num_voted = sum(voted_in_2022),              # Total number that voted
#     .groups = "drop"                             # Ensure the result is ungrouped
#   ) %>%
#   mutate(
#     pct_voted = round(100*(num_voted/total_respondents),2)
#   )
# 
# 
# ggplot(age_data, aes(x = age_bracket, fill = presvote2020)) +
#   geom_bar(aes(y = total_respondents, group = presvote2020), 
#            stat = "identity", position = "dodge", alpha = 0.5) +
#   geom_bar(aes(y = num_voted, group = presvote2020), 
#            stat = "identity", position = "dodge",  alpha = 1) +
#   # scale_fill_manual(values = c("Joe Biden" = "blue", "Donald Trump" = "red")) +
#   scale_fill_brewer(palette = "Set1") +
#   labs(
#     title = "Lower Turnout Across All Age Brackets and Vote Preferences in 2022",
#     x = "Age Bracket",
#     y = "Number of Respondents",
#     fill = "2020 Vote"
#   ) +
#   coord_flip() +
#   theme_minimal() +
#   theme(
#     legend.position = "top",
#     axis.text.x = element_text(angle = 45, hjust = 1)
#   )

# Graph the percentage of 2020 voters who also voted in 2022

age_data <- ces2022_analysis_data %>%
  group_by(age_bracket, presvote2020) %>%
  summarise(
    total_respondents = n(),                     # Total number of respondents
    num_voted = sum(voted_in_2022),              # Total number that voted
    .groups = "drop"                             # Ensure the result is ungrouped
  ) %>%
  mutate(
    pct_voted = round(100*(num_voted/total_respondents),2)
  )

# Plotting the graph
# Graph the percentage of 2020 voters who also voted in 2022

age_data <- ces2022_analysis_data %>%
  group_by(age_bracket, presvote2020) %>%
  summarise(
    total_respondents = n(),                     # Total number of respondents
    num_voted = sum(voted_in_2022),              # Total number that voted
    .groups = "drop"                             # Ensure the result is ungrouped
  ) %>%
  mutate(
    pct_voted = round(100*(num_voted/total_respondents),2)
  )

# Plotting the graph
ggplot(age_data, aes(x = age_bracket, fill = presvote2020)) +
  geom_bar(aes(y = total_respondents, group = presvote2020), 
           stat = "identity", position = "dodge", alpha = 0.5) +  # Lower alpha for non-voters
  geom_bar(aes(y = num_voted, group = presvote2020), 
           stat = "identity", position = "dodge",  alpha = 1) +  # Higher alpha for voters
    scale_fill_brewer(palette = "Set1") +
  labs(
    title = "Lower Turnout Across All Age Brackets and Vote Preferences in 2022 than 2020",
    x = "Age Bracket",
    y = "Number of Respondents"
  ) +
  coord_flip() +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(size = 12) 
  ) +
  scale_alpha_manual(
    values = c(0.5, 1),
    name = "Voter Turnout",
    labels = c("Did Not Vote in 2022", "Voted in 2022")
  )
```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

# trust in the federal government
interest_data <- ces2022_analysis_data %>%
  group_by(political_interest, presvote2020) %>%
  summarise(
    total_respondents = n(),                     # Total number of respondents
    num_voted = sum(voted_in_2022),              # Total number that voted
    .groups = "drop"                             # Ensure the result is ungrouped
  ) %>%
  mutate(
    pct_voted = round(100*(num_voted/total_respondents),2),
    prop_voted = round(num_voted/total_respondents, 2)
  )

interest_data <- interest_data %>%
  mutate(political_interest = factor(political_interest, 
                                     levels = c("No", "Hardly at all", "Only now and then", "Some of the time", "Most of the time")))


ggplot(interest_data, aes(x = political_interest, y = prop_voted, fill = presvote2020)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +  # Use dodge position for side-by-side bars
 scale_fill_brewer(palette = "Set1") + # Customize colors for Biden and Trump
  coord_flip() +
  labs(
    title = "2020 Voters with Lower Political Interest Less Likely to Vote in 2022",
    x = "Political Interest",
    y = "% of 2020 Voters who Voted in 2022 Midterms",
    fill = "2020 Presidential Vote"
  ) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +  # Set the scale from 0% to 100%
  theme_minimal() +  # Clean minimal theme
  theme(
    legend.position = "right",  # Move the legend to the top
    axis.text.x = element_text(angle = 45, hjust = 1, size=9),  # Rotate x-axis labels for better readability
    plot.title = element_text(size = 10)  # Adjust the title font size
  )
```
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-civicknowledge
#| fig-cap: There is a nearly 15-point gap in 2022 turnout amongst respondents who voted for Trump or Biden in 2020 and know which party has a majority in the U.S. House of Representatives versus those who do not. 

knowledge_data <- ces2022_analysis_data %>%
  group_by(know_us_house, presvote2020) %>%
    summarise(
    total_respondents = n(),                     # Total number of respondents
    num_voted = sum(voted_in_2022),              # Total number that voted
    .groups = "drop"                             # Ensure the result is ungrouped
  ) %>%
  mutate(
    pct_voted = round(100*(num_voted/total_respondents),2),
    prop_voted = round(num_voted/total_respondents, 2)
  )

ggplot(knowledge_data, aes(x = know_us_house, y = prop_voted, fill = presvote2020)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +  # Use dodge position for side-by-side bars
 scale_fill_brewer(palette = "Set1") + # Customize colors for Biden and Trump
  coord_flip() +
  labs(
    title = "2020 voters with higher civic knowledge more likely to vote in 2022 midterms",
    x = "Knows which party controls the House",
    y = "% of 2020 Voters who Voted in 2022 Midterms",
    fill = "2020 Presidential Vote"
  ) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +  # Set the scale from 0% to 100%
  theme_minimal() +  # Clean minimal theme
  theme(
    legend.position = "bottom",  # Move the legend to the top
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    plot.title = element_text(size = 12)  # Adjust the title font size
  )
```



# Model {#sec-model}
The goal of my modeling strategy is to use logistic regression models to investigate ...
I use two logistic regression models: one to model the probability that an individual who voted in the 2020 presidential election would also vote in the 2022 midterm election, and one to model the probability that an individual voter supports Trump.

The goal of my modeling strategy is twofold. Firstly, I want to The goal of our modelling strategy is twofold. Firstly,...

The model to predict $voted\_in\_2022$ uses the following predictors:

The model to predict $voted\_for\_trump$ uses the following predictors:

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

### Turnout model
Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

Models the probability that someone who voted for either Trump or Biden in 2020 also voted in the 2022 mid-term elections.

Define $\pi_i$ as the probability that survey respondent $i$ voted in the 2022 midterm election
```{=tex}
\begin{align*} 
y_i | \pi_i &\sim \mbox{Bern}(\pi_i) \\
\mbox{logit}(\pi_i) &= \beta_0 + \beta_1 \times presvote2020_i + \beta_2 \times educ_i + \beta_3 \times trust_i + \beta_4 \times know\_us\_house_i \\
&+ \beta_5 \times know\_us\_senate_i + \beta_6 \times political\_interest_i  \\
\beta_0 &\sim ~ \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim ~ \mbox{Normal}(0, 2.5) \\
\beta_2 &\sim ~ \mbox{Normal}(0, 2.5) \\
\beta_3 &\sim ~ \mbox{Normal}(0, 2.5) \\
\beta_4 &\sim ~ \mbox{Normal}(0, 2.5) \\
\beta_5 &\sim ~ \mbox{Normal}(0, 2.5) \\
\beta_6 &\sim ~ \mbox{Normal}(0, 2.5) \\
\end{align*}
```

Models the probability that a respondent voted for Trump in 2020:
```{=tex}
\begin{align*} 
y_i | \pi_i &\sim \mbox{Bern}(\pi_i) \\
\mbox{logit}(\pi_i) &= \beta_0 + \beta_1 \times age\_bracket_i + \beta_2 \times gender_i + \beta_3 \times race_i + \beta_4 \times educ\_i \\
&+ \beta_5 \times trust_i + \beta_6 \times know\_us\_house-i  \\
\beta_0 &\sim ~ \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim ~ \mbox{Normal}(0, 2.5) \\
\beta_2 &\sim ~ \mbox{Normal}(0, 2.5) \\
\beta_3 &\sim ~ \mbox{Normal}(0, 2.5) \\
\beta_4 &\sim ~ \mbox{Normal}(0, 2.5) \\
\beta_5 &\sim ~ \mbox{Normal}(0, 2.5) \\
\beta_6 &\sim ~ \mbox{Normal}(0, 2.5) \\
\end{align*}
```
We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.
Confounders have not been eliminated - 

I found that trust in the federal government is related to both support for Trump and likelihood that someone who voted in 2020 also voted in 2022. education same thing.

## Model Weaknesses and Limitations

Trained using a random subset of the data. Refer to codebook about the risks of that.
I randomly sampled 5,000 observations from the 2022 CES survey data set. This was a tradeoff; training a model on more observations would have had a significantly longer runtime. However, randomly subsetting the 2022 CES survey data set is risky because ...

# Results {#sec-resuls}

The results are summarized in [Appendix -@sec-model-results].

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

vote_preference_model <- 
  readRDS(file = here::here("models/political_preference_model.rds"))

turnout_model_2022 <-
  readRDS(file = here::here("models/turnout_model_2022.rds"))
```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

# # apply the model to predict who would vote in 2022
# who_voted <- ces2022_analysis_data %>%
#   group_by(voted_in_2022, presvote2020, age_bracket, educ, truststate, trustfed, 
#          know_us_house, know_us_senate, political_interest) %>%
#   summarise(n = n()) %>%
#   ungroup() 
# 
# # maybe look at young men, no collegee education, little trust, low informaton, low political interest and their support for trump and turnout when he was and was not on the ballot
# 
# # apply model to the analysis data
# # use predict function to predict the probability that each voter will vote based on characteristics
new_voter_data <- ces2022_analysis_data
new_voter_data$predicted_prob <- predict(turnout_model_2022, 
                                         newdata = new_voter_data, 
                                         type = "response")

new_voter_data <- new_voter_data |>
  select(predicted_prob, voted_in_2022, presvote2020, gender, age_bracket, educ,
         truststate, trustfed, know_us_house, political_interest)

turnout_model_results <- new_voter_data %>%
  group_by(presvote2020, age_bracket, educ, trustfed, truststate, know_us_house, political_interest) %>%
  summarise(turnout_estimate = mean(predicted_prob),
            turnout_actual = mean(voted_in_2022))

# # pretty good at classifying 18-29 year old trump voters who did not vote in midterms correctly
turnout_model_results_young_men <- new_voter_data %>%
  filter(gender == "Male" & age_bracket == "18-29")
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| eval: true

vote_preferences <- ces2022_analysis_data %>%
  group_by(age_bracket, gender, race, educ, trustfed, know_us_house) %>%
  summarise(trump_prop = mean(voted_for_trump))

new_preference_data <- ces2022_analysis_data
new_preference_data$trump_prob <- predict(vote_preference_model,
                                          newdata=new_preference_data,
                                          type="response")

new_preference_data <- new_preference_data|>
  select(trump_prob, voted_for_trump, presvote2020, age_bracket, gender, 
                              race, educ, trustfed, know_us_house)

preference_results <- new_preference_data %>%
  group_by(age_bracket, gender, race, educ, trustfed, know_us_house) %>%
  summarise(trump_estimate = mean(trump_prob),
            trump_actual = mean(voted_for_trump))
```

## Trump support amongst young men
```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false

trump_young_men <- new_preference_data |>
  filter(gender == "Male" & age_bracket == "18-29")

turnout_young_men <- new_voter_data |>
  filter(gender == "Male" & age_bracket == "18-29")

# Calculate proportions
trump_young_men <- trump_young_men |>
  mutate(predicted_vote_choice = if_else(trump_prob >= 0.5, "Trump", "Biden")) |>
  group_by(educ, predicted_vote_choice) |>
  summarise(count = n(), .groups = "drop") |>
  group_by(educ) |>
  mutate(percentage = count / sum(count) * 100)

# Create the plot with percentages
trump_young_men |> 
  ggplot(aes(x = educ, y = percentage, fill = predicted_vote_choice)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  labs(
    x = "Highest education",
    y = "Estimated % of voters",
    fill = "Voted for"
  ) +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

## Low Propensity Trump Voters are Less Likely to Vote in Midterms
This result foreshadowed the fact that young men, especially those without a college degree who are infrequent voters, would become a key part of Trump's winning coalition in 2024.
```{r}
# young men who voted for trump, no college education, low trust, do not know us house
young_men_lp <- turnout_young_men %>%
  filter(presvote2020 == "Donald Trump" &
           (educ == "High school graduate" | educ == "No HS" | educ == "Some college") &
           (trustfed == "None at all" | trustfed == "Not very much") &
           know_us_house == "No"
         )

young_men_lp_trump <- new_preference_data |>
  filter(gender == "Male" & age_bracket == "18-29") %>%
    filter( (educ == "High school graduate" | educ == "No HS" | educ == "Some college") &
           (trustfed == "None at all" | trustfed == "Not very much") &
           know_us_house == "No"
         )

```

## Voters with Little Interest in Politics are Less Likely to Vote in Midterms

## College Education is Associated with Higher Support for the Democratic Party and Higher Likelihood to Vote in Midterms

## Young men are more likely to support Trump and less likely to vote when he is not on the ballot

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Polling Misses and Trump's Unique Appeal to Low Propensity Voters
Young men who are not college-educated, do not know which party controls the U.S. House of Representatives, 

## 2024 Polling Misses and Non-Response Bias

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Surveys, Sampling, and Observational Data

## Deep Dive into CES Methodology

### Sampling Method

### Weighting

### Verification

### Weaknesses and Limitations

## Second Point: Idealized Survey Methodology to Study 2024 Down-Ballot Voter Attrition

## Social Desirability Bias and Identifying Non-Voters

# Additional data details

## Data cleaning

# Model details {#sec-model-details}

## Model Results {#sec-model-results}

The model summary is shown in @tbl-model-summary.

```{r}
#| echo: false
#| eval: true
#| label: tbl-model-summary
#| tbl-cap: "Explaining whether someone who voted in the 2020 presidential election also voted in the 2022 midterm election, based on who they voted for, age, education, trust in government, civic knowledge, and political interest "
#| warning: false

modelsummary::modelsummary(
  list(
    "Voter Turnout Model" = turnout_model_2022
  ),
  statistic = "mad",
  fmt = 2,
  output = "kableExtra"
) %>%
  kableExtra::kable_styling(font_size = 11)
```
@tbl-modelsummary makes

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows the comparison of the outcome variable $vote\_2022$, with simulations from the posterior distribution [@tswd].

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

# pp_check(turnout_model_2022) +
#   theme_classic() +
#   theme(legend.position = "bottom")
# 
# posterior_vs_prior(turnout_model_2022) +
#   theme_minimal() +
#   scale_color_brewer(palette = "Set1") +
#   theme(legend.position = "bottom") +
#   coord_flip()
```

## Credibility Intervals

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-credibilityintervals
#| fig-cap: Credible intervals for predictors of voter turnout in the 2022 U.S. midterm elections

modelplot(turnout_model_2022, conf_level = 0.95) +
  labs(x = "95 per cent credibility interval")

```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(turnout_model_2022, "trace")

plot(turnout_model_2022, "rhat")
```

\newpage

# References
